ms2

### 2017-08-16 ms2/ms2_display.Rmd
new figures; mup

```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
ineq = read_tsv('ineq.tsv') %>%
  group_by(drug_group, unit_type, unit) %>%
  summarize_at(vars(mean, fnz), mean) %>%
  ungroup()

@@

```{r}
results %>%
  filter(model=='univariate_mean', unit_type=='hrr', term=='mean') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, HRR")
```

@@

@@

We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$. We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.

These plots show the coefficient $\beta_\mu$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu} \mu^{(u)} + \varepsilon^{(u)}$.

```{r hrr_multi}
results %>%
  filter(model=='multivariate', unit_type=='hrr', term=='mean') %>%
  interval_plot() +
  ylab(expression(beta[mu])) +
  ggtitle("Linear multivariate, HRR")
```

```{r state_multi}
results %>%
  filter(model=='multivariate', unit_type=='state', term=='mean') %>%
  interval_plot() +
  ylab(expression(beta[mu])) +
  ggtitle("Linear multivariate, state")
```

@@

The second two, *E. faecium* with $\beta$-lactams and pneumococcus with tetracyclines, are not problematic: in both cases, neither $R \sim f_>$ nor $R \sim \mu$ is significant. We're only interested in cases where $f_>$ gives information and $\mu$ gives information beyond that. There are cases where $f_>$ doesn't appear to give any information.

@@

First, the regression results are equivocal. Regressing $r ~ \mu$, where $r$ are the residuals from the regression $R ~ f_>$, gives a non-significant result. An ANOVA analysis shows that sequential application of $f_>$ and $\mu$ as predictors of $R$ both explain non-trivial variance; but if the order is reversed, and $\mu$ is used first, then it *doesn't* explain any variance. This makes me think that there is just some weird interaction between the two terms in the multiple regression.

@@

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim resistance E. coli') +
  ggtitle('E. coli & Bactrim across states')
```

@@

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. coli') +
  ggtitle('E. coli & beta-lactams across states')
```

>>>>

```{r load_data, cache.extra=c(md5sum('ineq.tsv'), md5sum('results.tsv'))}
ineq = read_tsv('ineq.tsv') %>%
  group_by(drug_group, unit_type, unit) %>%
  summarize_at(vars(mean, fnz), mean) %>%
  ungroup() %>%
  mutate(mup=mean/fnz)

@@

@@

Rather than mean consumption $\mu$, I'll use the mean consumption among consumers $\mu_> = \mu / f_>$.

```{r fnz_mup}
ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mup, color=drug_group)) +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab(expression(f['>'])) +
  ylab(expression(mu['>'])) +
  ggtitle('Mean nonzero consumption vs. fraction consuming, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

The tetracyclines and Bactrim are the interesting exceptions: in these drugs, states where consumption is more widespread are the states where consumption per consumer is lower.

@@

We ask if mean nonzero consumption $\mu_>$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$. We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.

These plots show the coefficient $\beta_{\mu_>}$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu_>} \mu_>^{(u)} + \varepsilon^{(u)}$.

```{r hrr_multi}
results %>%
  filter(model=='multivariate', unit_type=='hrr', term=='mup') %>%
  interval_plot() +
  ylab(expression(beta[mu['>']])) +
  ggtitle("Linear multivariate, HRR")
```

At the state level, none of the $\beta_{\mu_>}$ coefficients are significantly different from zero (after multiple-hypothesis correction).

```{r state_multi}
results %>%
  filter(model=='multivariate', unit_type=='state', term=='mup') %>%
  interval_plot() +
  ylab(expression(beta[mu['>']])) +
  ggtitle("Linear multivariate, state")
```

@@

For *E. faecium* with $\beta$-lactams, the univariate regression with $f_>$ is not significant, and the one with $\mu_>$ is marginally so. Similarly, for *S. pneumoniae* and tetracyclines, the regression with $f_>$ is not significant, and the regression wiht $\mu_>$ is. Thus, it looks like $\mu_>$ isn't adding information in addition to $f_>$; instead, it just has all the information.

@@

**swo double-check** First, the regression results are equivocal. Regressing $r ~ \mu$, where $r$ are the residuals from the regression $R ~ f_>$, gives a non-significant result. An ANOVA analysis shows that sequential application of $f_>$ and $\mu$ as predictors of $R$ both explain non-trivial variance; but if the order is reversed, and $\mu$ is used first, then it *doesn't* explain any variance. This makes me think that there is just some weird interaction between the two terms in the multiple regression.

@@

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim resistance E. coli') +
  ggtitle('E. coli & Bactrim across states')
```

### E. faecium and beta-lactams

```{r efaecium_betalactam}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance in E. faecium') +
  ggtitle('E. faecium & beta-lactams across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. coli') +
  ggtitle('E. faecium & beta-lactams across states')
```

### S. pneumonia and tetracyclines

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed tetra') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed tetra') +
  ylab('tetra resistance pneumo') +
  ggtitle('pneumo & tetra across states')
```

@@

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='beta_lactam') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy point
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. coli') +
  ggtitle('E. coli & beta-lactams across states')
```

## Comparing the predictive power of mean consumption and fraction nonzero

Not surprisingly, linear regressions $R \sim \mu_>$ give slightly different results from $R \sim f_>$. The negative correlations are all in those drugs where $\mu_>$ was anticorrelated with $f_>$.

```{r hrr_linear_mup}
results %>%
  filter(model=='univariate_mup', unit_type=='hrr', term=='mup') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[mu['+']])) +
  ggtitle("Linear univariate, HRR")
```

How do the regressions $R \sim f_>$ compare to $R \sim \mu_>$? Here I look at the correlation between the coefficients $\beta_{\mu_>}$ and $\beta_{f_>}$ across bug/drug combinations. Not surprisingly, $\beta_\mu$ and $\beta_{f_>}$ are highly correlated: because $\mu$ and $f_>$ are correlated, you can guess the result of the one regression from the other. However, $\beta_{\mu_>}$ is not as well correlated with $\beta_{f_>}$, indicating that there might be some extra information there.

```{r univariate_result_correlation}
results %>%
  filter(str_detect(model, '^uni'), term != '(Intercept)') %>%
  select(unit_type, bug, drug_group, term, estimate) %>%
  spread(term, estimate) %>%
  group_by(unit_type, drug_group) %>%
  filter(n() > 2) %>%
  do(tidy(cor(select(., -unit_type, -bug, -drug_group)))) %>%
  ungroup() %>%
  filter(!is.na(fnz)) %>%
  rename(x=.rownames) %>%
  gather('y', 'value', fnz:mup) %>%
  filter(x > y) %>%
  mutate(key=str_c(x, '_', y)) %>%
  select(-x, -y) %>%
  spread(key, value) %>%
  kable
```

In other words, if you do the regression with $f_>$, you can predict with very high accuracy the coefficient from the univariate regression on $\mu$.
